The Symmetry of Partner Modelling

### 2016-04-18 mutual-modeling.tex
Integrated proof-reading from CSCL editors

\usepackage{hyperref}
\usepackage{subcaption}

@@

Collaborative learning has often been associated with the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modelling, \ie that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term ``some
model'' to stress the fact that this model is not necessarily detailed or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner?
Upon the modelee's ability to make his state clear to the modeler? Or rather upon
the quality of their interactions? We address this question by comparing
the respective accuracies of the models built by different team
members. We report on 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that \A built about
\B was correlated with the accuracy of the model that \B built about \A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these findings do not
rule out the fact that individual abilities also participate in the quality of
modelling process.

@@

From its inception, Computer Supported Collaborative Learning (CSCL)
research has been following the suggestion by \citet{roschelle1995construction}
that collaborative learning has something to do with the process of constructing
and maintaining a \emph{shared understanding} of the task at hand. Building a
shared/mutual understanding refers to the upper class of collaborative learning
situations, those in which students should build upon each other's understanding
to refine their own understanding: what is expected to produce learning is not
the mere fact that two students build the same understanding but the cognitive
effort they have to engage to build this shared
understanding~\citep{schwartz1995emergence}. This effort can be observed by the
frequency of rich interactions, \ie interactions whose occurrence has been
related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social} and mutual
regulation~\citep{blaye1995collaborative} in a Vygostkian perspective. The
construction of a shared understanding has been investigated for several decades
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in the learning
sciences. Grounding mechanisms are appropriate to explain conversational
events, such as referential failures in short dialogue episodes, but they hardly
predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
acceptation and integration of a new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

\begin{quote}
Collaborative learning is associated with the increased
cognitive-interactional effort involved in the transition from \emph{learning to
understand each other} to \emph{learning to understand the meanings of the semiotic
tools that constitute the mediators of interpersonal
interaction}~\citep{baker1999role}
\end{quote}

Along this line, several scholars suggest that CSCL research should go deeper in
the understanding of how partners engage into shared meaning
making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
making~\citep{suthers2006technology}.

Paradoxically, while Clark's theory is somewhat too linguistic from a learning
viewpoint, it is criticized at the same time as being too cognitivist by some
psycholinguists, \ie as overestimating the amount of shared knowledge and mutual
representations actually necessary to conduct a dialogue. The fundamental issue,
as old as philosophy, is the degree of coupling between the different levels of
dialogue, mostly between the lexical/syntactical level and the deeper semantic
levels. \citet{pickering2006alignment} argue that mutual understanding
starts mostly with a \emph{superficial alignment} at the level of the linguistic
representations, due to priming mechanisms, and that this local alignment may --
in some cases -- lead to a \emph{global alignment} of the semantic level
(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
even the repair of some misunderstandings, is explained by this mimetic behavior
more than by a monitoring of each other's knowledge: \emph{``...interlocutors do
not need to monitor and develop full common ground as a regular, constant
part of routine conversation, as it would be unnecessary and far too costly.
Establishment of full common ground is, we argue, a specialized and
non-automatic process that is used primarily in times of difficulty (when
radical misalignment becomes apparent).''} (p. 179).  This view is actually not
incompatible with Clark's \emph{grounding
criterion}~\citep{clark1989contributing}: the degree of shared understanding
that peers need to reach depends upon the task they perform. For instance, a
dialogue between two surgeons might rely on superficial alignment if they talk
about their friends but has to guarantee accurate common grounds when talking
about which intervention will be conducted in which way on which patient.  In
this paper, we operationalized the grounding criterion, \ie the necessity for
accurate modelling, as \emph{the correlation between the accuracy of partner models
and measures of team performance}.

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: Is linguistic alignment sufficient to trigger conceptual
change? Does negotiation of meaning only occur when partners monitor
and diagnose each other's knowledge? If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with simple referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In the studies we report here, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall \emph{least collaborative
effort}.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings, and in fact,
subjects tend to make less efforts adapting their utterances to a specific
partner when they know that they can later provide feedback on his/her

understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts}~\citep{kobbe2007specifying}. A script is a pedagogical
method that frames collaborative learning activities in order to foster the
emergence of productive interactions such as argumentation, explanation or
conflict. Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} form pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, {\sc
jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
complementary knowledge for augmenting (reasonably) the efforts that group members
have to engage into to reach a shared solution.

@@

We refer to \emph{Partner modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mental modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling.
Collaborative learning dialogue include many instances of utterances such as ``I
thought he would do that'' (first order modelling) or even ``He
thought I would do that but I intended something else.'' (second order
modelling).

@@

Mutual modelling is bi-directional. During dyadic problem solving,   partner A
builds some model of \B and \B build some model of \A.  Moreover, these two
processes are not independent: \A's model of what \B knows, include what \B knows
about \A.  This leads to nested levels of  modelling. If \A states ``\B thinks I
am good in maths'', \A builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: \A saying ``\B knows that I don't expect him to
solve this statistics problem'' corresponds to \\
\model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
one study we report on, mutuality also applies to triads in which \A will elaborate
a model of \B and of C, and reciprocally. This is probably also true for larger
groups although one may hypothesize that there exists (yet undefined) group size
from which it is not possible to model all partners individually and, therefore,
members  model  the group as a whole instead as a collection of individuals.  We
do not explore this hypothesis here and limit ourselves to dyads and triads.

@@

Natural language becomes cumbersome when describing things such as ``the model
that \A builds about the model that \B builds about A''. Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean \A has an
explicit, monolithic representation of \B: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, the model built by \A can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; its usefulness is internal to this paper.  Besides, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university (and they hence had some general common
        grounds).  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a {\sc jigsaw} script.
        \vspace{1em}

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provides \A with information about
        what \B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

The experiments we report here address mutual modelling across different tasks,
some with dyads, other with triads. They were conducted along 6 years in two
different institutions by different researchers. They used different
independent, intermediate and dependent variables. Nonetheless, we were able to
retroactively address 3 research questions across these studies.

@@

\begin{table*}[ht!]
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\centering
%\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.1cm}L{1.65cm}L{1.7cm}L{1.8cm}L{1.6cm}L{1.6cm}}
\toprule
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ 
\midrule
\textbf{Task}                 & game in virtual space            & game in physical space          & building an argument map  & building a concept map    & building a concept map       \\
\midrule
\textbf{Interactions}         & audio                            & written                         & audio / video             & audio                     & audio                        \\
\midrule
\textbf{Shared editor}        & 3D space                         & 2D map                          & concept map               & concept map               & concept map                  \\
\midrule
\textbf{Group size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\midrule
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\midrule
\textbf{Awareness tool}       & partner's position               & partner's current/past pos.     & -                         & partner's concept map     & partner's scores at pre-test \\
\midrule
\textbf{Dependent variable}       & team/indiv. performance            & team/indiv. performance           & team performance          & team/indiv. knowledge       & team/indiv. knowledge          \\
\midrule
\textbf{Independent variable}     & awareness tool \textit{vs} none  & awareness tool \textit{vs} none & audio \textit{vs} audio+video & script \textit{vs} none & awareness tool \textit{vs} none       \\
%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\bottomrule
\end{tabular}
%}
\caption{Overview of the studies}
\label{synthesis_table}
\end{table*}

@@

Thirty-six persons participated in this study, all native French speakers. We
constituted 18 dyads who did not know each other beforehand. The pairs were
randomly assigned to either the control condition (without the awareness tool)
or the awareness condition (with the awareness tool).

@@

How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is
nonetheless far from 1, which implies we cannot discard the individual modelling
skills. We collect more evidence in the next studies.

@@

This study concerns a collaborative game that occurred in the physical
space~\citep{nova2006underwhelming}. We studied whether players
build an accurate model of the path followed by their partners, assuming that this
path would reflect their problem solving strategy. We used an objective measure of
\gModel{A}{B}: the distance between where \A believes \B has been walking and
where \B actually went.  The main hypothesis concerned the effect of awareness
tools on group performance and on \gModel{A}{B}.

@@

{\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
joint task. The game was played on our campus. Participants had to find a
virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
The players used a Tablet PC that displayed a map of the campus and an indication of
their personal distance to Bob. Their annotations on the map were shared with
the two other players (\A could see what \B and \C wrote). These annotations
faded out after a few minutes to avoid covering the full display. The awareness
tool displayed as well the location of the two other players on the map. Three
conditions were considered: the control condition (without tool) and two
experimental conditions: synchronous awareness (display of the current position
of each player) and asynchronous awareness (display of current position of each
player as well as their recent spatial trace).

@@

Ninety students participated in this experiment. We only selected students from
our campus since knowledge of the campus geography had an impact both on group
performance and on mutual modelling: to represent the path of someone across
some space is difficult without an \textit{a priori} mental map of this space.
We formed groups of students who knew each other. We assigned 10 triads to each
of our three experimental conditions. Each condition was made up of
approximately 25\% of women, but we did not control gender repartition within
each triad.

@@

The aim of this unpublished\footnote{The data and statistical analyses of this
study are available online:
\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.} study was
to evaluate the effect of media richness on \Model{A}{B}{emotions}.  The
hypothesis was that video communication would lead to a better \gModel{A}{B}
than audio only since emotions often impact facial expressions.

@@

This study investigated the effect of a collaboration script on collaborative
learning~\citep{molinari2008effects}. The script chosen is a
{\sc jigsaw}: two students receive different but complementary subsets of the
knowledge (texts) which have to be integrated to build a shared concept map.
This script increases the cognitive effort to build the map, not only to
conciliate the viewpoints of each team members but, before that, to find out
what the other knows.

@@

Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
were remunerated for participation. Dyads were randomly assigned to one of the
two experimental conditions. Gender was balanced over the conditions.
Participants did not know each other before the experiment. Students from Life
Sciences faculty were not recruited to avoid high initial background knowledge on the
learning domain.

@@

\paragraph{Grounding criterion} In this task, the grounding criterion was low.
We did not however evaluate task performance (\eg the quality of the jointly
produced concept map) but learning gains. The correlation between \gModel{A}{B}
and \A's learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
also not significant within each condition.

@@

\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
0.05, ns$), which does not support $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. In
addition, the KAT supports the processing of modelling at the first level
($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$). Hence the
fact that the KAT enhances  \gModel{A}{B}  brings additional supports to
$\mathcal{H}_{1}$. However, the analysis of verbal interactions finds elements
that actually supports the three hypotheses.

@@

In the introduction, we mentioned the controversy around the cognitive depth of
dialogues: does efficient dialogue require some modeling of what the partner
understands or intends to convey, as initially postulated by
\citet{clark1986referring}, or can dialogue simply rely on some shallow
syntactic alignment, as objected by \citet{pickering2006alignment}. While the
shallow hypothesis may be relevant to simple chat situations, we investigated
this issue in richer problem solving tasks that are more representative of the
tasks assigned to learners in CSCL environments. In four different tasks, we
found evidence that partner model each other since the quality of modelling
emerged as an intermediate variable, sensitive to several independent variables
(like the presence of awareness tools or media richness), and predictive of
several dependent variables, such as task performance or learning gains.
Thereby, even if we do not bring any definite conclusion to this debate, our
results support the CSCL school of thought in which shared understanding or
intersubjective meaning making have been a foundational concept
\citep{roschelle1995construction,schwartz1995emergence,dillenbourg2006sharing,suthers2006technology,stahl2007meaning}.

>>>>

\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{endnotes}

@@

Collaborative learning has often been associated with the construction of  a
shared   understanding of the situation at hand. The psycholinguistics
mechanisms at work while establishing common grounds are the object of
scientific controversy. We postulate that collaborative tasks require some
level of mutual modelling, \ie that each partner needs some model of what
the other partners know/want/intend at a given time.  We use the term ``some
model'' to stress the fact that this model is not necessarily detailed or
complete, but that we acquire some representations of the persons we
interact with. The question we address is: Does the quality of the partner
model depend upon the modeler's ability to represent his or her partner?
Upon the modelee's ability to make his state clear to the modeler? Or rather, upon
the quality of their interactions? We address this question by comparing
the respective accuracies of the models built by different team
members. We report on 5 experiments on collaborative problem solving
or collaborative learning that vary in terms of tasks (how important it is
to build an accurate model) and settings (how difficult it is to build an
accurate model). In 4 studies, the accuracy of the model that \A built about
\B was correlated with the accuracy of the model that \B built about \A,  which
seems to imply that the quality of interactions matters more than individual
abilities when building mutual models.  However, these findings do not
rule out the fact that individual abilities also contribute to the quality of
modelling process.

@@

From its inception, Computer Supported Collaborative Learning (CSCL)
research has been following the suggestion by \citet{roschelle1995construction}
that collaborative learning has something to do with the process of constructing
and maintaining a \emph{shared understanding} of the task at hand. Building a
shared/mutual understanding refers to the upper class of collaborative learning
situations, those in which students should build upon each other's understanding
to refine their own understanding: what is expected to produce learning is not
the mere fact that two students build the same understanding but the cognitive
effort they have to engage to build this shared
understanding~\citep{schwartz1995emergence}. This effort can be observed by the
frequency of rich interactions, \ie interactions whose occurrence has been
related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social} and mutual
regulation~\citep{blaye1995collaborative} in a Vygostkian perspective. The
construction of shared understanding has been investigated for several decades
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in the learning
sciences. Grounding mechanisms are appropriate to explain conversational
events, such as referential failures in short dialogue episodes, but they hardly
predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
acceptance and integration of a new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

\begin{quote}
Collaborative learning is associated with the increased
cognitive-interactional effort involved in the transition from \emph{learning to
understand each other} to \emph{learning to understand the meanings of the semiotic
tools that constitute the mediators of interpersonal
interaction}~\citep[p.31]{baker1999role}
\end{quote}

\noindent Along this line, several scholars suggest that CSCL research should go deeper 
towards understanding how partners engage in shared meaning
making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
making~\citep{suthers2006technology}.

Paradoxically, while Clark's theory is somewhat too linguistic from a learning
viewpoint, it is criticized at the same time as being too cognitivist by some
psycholinguists, \ie as overestimating the amount of shared knowledge and mutual
representations actually necessary to conduct a dialogue. The fundamental issue,
as old as philosophy, is the degree of coupling between the different levels of
dialogue, mostly between the lexical/syntactical level and the deeper semantic
levels. \citet{pickering2006alignment} argue that mutual understanding
starts mostly with a \emph{superficial alignment} at the level of the linguistic
representations, due to priming mechanisms, and that this local alignment may --
in some cases -- lead to a \emph{global alignment} of the semantic level
(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
even the repair of some misunderstandings, is explained by this mimetic behavior
more than by a monitoring of each other's knowledge:

\begin{quote}
[\ldots] interlocutors do
not need to monitor and develop full common ground as a regular, constant
part of routine conversation, as it would be unnecessary and far too costly.
Establishment of full common ground is, we argue, a specialized and
non-automatic process that is used primarily in times of difficulty (when
radical misalignment becomes apparent).~\citep[p.179]{pickering2006alignment}
\end{quote}

\noindent This view is actually not
incompatible with Clark's \emph{grounding
criterion}~\citep{clark1989contributing}: the degree of shared understanding
that peers need to reach depends upon the task they perform. For instance, a
dialogue between two surgeons might rely on superficial alignment if they talk
about their friends but has to guarantee accurate common grounds when talking
about which intervention will be conducted in which way on which patient.  In
this paper, we operationalized the grounding criterion, \ie the necessity for
accurate modelling, as \emph{the correlation between the accuracy of partner models
and measures of team performance}.

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: Is linguistic alignment sufficient to trigger conceptual
change? Does negotiation of meaning only occur when partners monitor
and diagnose each other's knowledge? If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with simple referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In the studies we report here, we
explore several richer tasks such as arguing about a sensitive issue or building a
concept map.

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall \emph{least collaborative
effort}.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings, and in fact,
subjects tend to make less efforts adapting their utterances to a specific
partner when they know that they can later provide feedback on his/her
understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts}~\citep{kobbe2007specifying}. A script is a pedagogical
method that frames collaborative learning activities in order to foster the
emergence of productive interactions such as argumentation, explanation or
conflict. Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} form pairs of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, and more load. Similarly, {\sc
jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
complementary knowledge for augmenting (reasonably) the efforts that group members
have to engage into to reach a shared solution.

@@

We refer to \emph{Partner modelling} as the process of inferring one's partner's
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mental modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling.
Collaborative learning dialogue include many instances of utterances such as ``I
thought he would do that'' (first order modelling) or even ``He
thought I would do that but I intended something else.'' (second order
modelling).

@@

Mutual modelling is bi-directional. During dyadic problem solving,   partner A
builds some model of \B and \B build some model of \A.  Moreover, these two
processes are not independent: \A's model of what \B knows, includes what \B knows
about \A.  This leads to nested levels of  modelling. If \A states ``\B thinks I
am good in maths'', \A builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: \A saying ``\B knows that I don't expect him to
solve this statistics problem'' corresponds to \\
\model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
one study we report on, mutuality also applies to triads in which \A will elaborate
a model of \B and of C, and reciprocally. This is probably also true for larger
groups although one may hypothesize that there exists (yet undefined) group size
from which it is not possible to model all partners individually and, therefore,
members  model  the group as a whole instead as a collection of individuals.  We
do not explore this hypothesis here and limit ourselves to dyads and triads.

@@

Natural language becomes cumbersome when describing things such as ``the model
that \A builds about the model that \B builds about A''. Therefore, to define
hypotheses and report on experiments on mutual modelling, we use  the notation
\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
proposed as a formal theory of mutual modelling but as useful simplification for
communicating about mutual modelling. This notation does not mean \A has an
explicit, monolithic representation of \B: it must be understood as an
abstraction referring to complex socio-cognitive processes.  As explained in the
previous section, the model built by \A can be fragmented, multi-dimensional, etc. This notation
is neither presented as a computational model of mutual modelling, nor as some
universal formalism; its usefulness is internal to this paper.  Additionally, we
refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
discuss in the next section the methodological difficulty in measuring this
accuracy.

@@

    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4, and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university (and they hence had some general common
        grounds).  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a {\sc jigsaw} script.
        \vspace{1em}

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1, 2, and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provides \A with information about
        what \B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

@@

The experiments we report here address mutual modelling across different tasks,
some with dyads, others with triads. They were conducted over 6 years in two
different institutions by different researchers. They used different
independent, intermediate, and dependent variables. Nonetheless, we were able to
retroactively address 3 research questions across these studies.

@@

\begin{table*}[ht!]
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\centering
%\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.1cm}L{1.65cm}L{1.7cm}L{1.8cm}L{1.6cm}L{1.6cm}}
\toprule
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ 
\midrule
\textbf{Task}                 & game in virtual space            & game in physical space          & building an argument map  & building a concept map    & building a concept map       \\
\midrule
\textbf{Interactions}         & audio                            & written                         & audio / video             & audio                     & audio                        \\
\midrule
\textbf{Shared editor}        & 3D space                         & 2D map                          & concept map               & concept map               & concept map                  \\
\midrule
\textbf{Group size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\midrule
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\midrule
\textbf{Awareness tool}       & partner's position               & partner's current/past pos.     & -                         & partner's concept map     & partner's scores at pre-test \\
\midrule
\textbf{Dependent variable}       & team/indiv. performance            & team/indiv. performance           & team performance          & team/indiv. knowledge       & team/indiv. knowledge          \\
\midrule
\textbf{Independent variable}     & awareness tool \textit{vs} none  & awareness tool \textit{vs} none & audio \textit{vs} audio+video & script \textit{vs} none & awareness tool \textit{vs} none       \\
%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\bottomrule
\end{tabular}
%}
\caption{Overview of the studies and main characteristics}
\label{synthesis_table}
\end{table*}

@@

Thirty-six persons participated in this study, all native French speakers. We
formed 18 dyads who did not know each other beforehand. The pairs were
randomly assigned to either the control condition (without the awareness tool)
or the awareness condition (with the awareness tool).

@@

How do we interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is
nonetheless far from 1, which implies we cannot discard the individual modelling
skills. We collect more evidence in the next studies.

@@

This study concerns a collaborative game that occurred in physical
space~\citep{nova2006underwhelming}. We studied whether players
build an accurate model of the path followed by their partners, assuming that this
path would reflect their problem solving strategy. We used an objective measure of
\gModel{A}{B}: the distance between where \A believes \B has been walking and
where \B actually went.  The main hypothesis concerned the effect of awareness
tools on group performance and on \gModel{A}{B}.

@@

{\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
joint task. The game was played on a university campus. Participants had to find a
virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
The players used a Tablet PC that displayed a map of the campus and an indication of
their personal distance to Bob. Their annotations on the map were shared with
the two other players (\A could see what \B and \C wrote). These annotations
faded out after a few minutes to avoid covering the full display. The awareness
tool also displayed the location of the two other players on the map. Three
conditions were considered: the control condition (without tool) and two
experimental conditions: synchronous awareness (display of the current position
of each player) and asynchronous awareness (display of current position of each
player as well as their recent spatial trace).

@@

Ninety students participated in this experiment. We only selected students from
university campus since knowledge of the campus geography had an impact both on group
performance and on mutual modelling: to represent the path of someone across
some space is difficult without an \textit{a priori} mental map of this space.
We formed groups of students who knew each other. We assigned 10 triads to each
of our three experimental conditions. Each condition was made up of
approximately 25\% of women, but we did not control gender repartition within
each triad.

@@

The aim of this unpublished\endnote{The data and statistical analyses of this
study are available online:
\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.} study was
to evaluate the effect of media richness on \Model{A}{B}{emotions}.  The
hypothesis was that video communication would lead to a better \gModel{A}{B}
than audio only since emotions often impact facial expressions.

@@

This study investigated the effect of a collaboration script on collaborative
learning~\citep{molinari2008effects}. The script chosen is a
{\sc jigsaw}: two students receive different but complementary subsets of the
knowledge (texts) which have to be integrated to build a shared concept map.
This script increases the cognitive effort to build the map, not only to
conciliate the viewpoints of each team member but, before that, to find out
what the other knows.

@@

Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
were remunerated for participation. Dyads were randomly assigned to one of the
two experimental conditions. Gender was balanced over the conditions.
Participants did not know each other before the experiment. Students from the Life
Sciences faculty were not recruited to avoid high initial background knowledge on the
learning domain.

@@

\paragraph{Grounding criterion} In this task, the grounding criterion was low.
We did not, however, evaluate task performance (\eg the quality of the jointly
produced concept map) but learning gains. The correlation between \gModel{A}{B}
and \A's learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
also not significant within each condition.

@@

\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
0.05, ns$), which does not support $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. In
addition, the KAT supports the processing of modelling at the first level
($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$). Hence the
fact that the KAT enhances  \gModel{A}{B}  brings additional supports to
$\mathcal{H}_{1}$. However, the analysis of verbal interactions finds elements
that actually support the three hypotheses.

@@

In the introduction, we mentioned the controversy around the cognitive depth of
dialogues: does efficient dialogue require some modeling of what the partner
understands or intends to convey, as initially postulated by
\citet{clark1986referring}, or can dialogue simply rely on some shallow
syntactic alignment, as objected by \citet{pickering2006alignment}. While the
shallow hypothesis may be relevant to simple chat situations, we investigated
this issue in richer problem solving tasks that are more representative of the
tasks assigned to learners in CSCL environments. In four different tasks, we
found evidence that partners model each other since the quality of modelling
emerged as an intermediate variable, sensitive to several independent variables
(like the presence of awareness tools or media richness), and predictive of
several dependent variables, such as task performance or learning gains.
Therefore, even if we do not bring any definite conclusion to this debate, our
results support the CSCL school of thought in which shared understanding or
intersubjective meaning making have been a foundational concept
\citep{roschelle1995construction,schwartz1995emergence,dillenbourg2006sharing,suthers2006technology,stahl2007meaning}.
